Timing and magnitude of climate driven range shifts in transboundary fish stocks challenge their management

### 2021-07-26 Scripts/Manuscript.Rmd
duplicated paragraph removed

Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (see Methods and Supplemental Materials). We then estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig. \@ref(fig:toeFigure1N)) and the intensity of this shift (i.e., time of emergence and stock share ratio; see Methods and Supplemental Materials). We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).

@@

<!-- We estimate the year in which 663 transboundary species (i.e., 9,132 stocks) previously identified by Palacios-Abrantes *et al.* (2020) [@PalaciosAbrantes:2020tv] would undergo changes in their shared distribution beyond historical variation and the intensity of said changes (see the Supplementary Materials). We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas). We defined the boundaries of the world’s EEZs using the Sea Around Us (http://www.seaaroundus.org/) spatial division, noting that it subdivides the EEZs of 198 coastal states into 280 regions (Fig. \@ref(fig:toeFigureS5)) including island territories (see the Supplementary Materials - *Databases and species selection*). -->

<!-- We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks under climate change (see the Supplementary Materials). In brief, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd]. The DBEM was driven with ten different simulations of the GFDL ESM2M (i.e., a ten-ensemble member simulation) each of them representing a unique atmosphere and ocean state at the start of the simulation [@Dunne:2012bo; @Dunne:2013db; @Frolicher:2020cc]. The simulations were run under historical forcing until 2005, afterwards and until 2100, simulations followed a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5) [@Riahi:2011dk]. -->

<!-- We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks. The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig. \@ref(fig:toeFigure1)A). We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution (see the Supplementary Materials - *Calculating shifts in the shared distribution of transboundary stocks*). The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ. \@ref(eq:four)). We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged). -->

<!-- Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (See Supplemental Material - *Quantifying the intensity of transboundary stocks’ range shift*). A threat point is defined as the minimum payoff a player would be willing to receive to cooperate in a game-theoretic model. Here, we define the minimum payoff as two standard deviations from the historical catch proportion (average of 1951 to 2005) that each EEZ had for each stock. This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp]. Finally, we computed changes in catch proportion that were below the threat point of neighboring EEZs sharing a stock for two time periods relative to the historical proportion: 2030 (2021 and 2040) matching the UN 2030 Agenda and a mid-century timeline (2040 to 2050). -->

@@

###mClimate risk hotspots for transboundary fisheries management

>>>>

Here, we employ a mechanistic population dynamic model driven by output from a comprehensive Earth system model with ten ensemble members to project transboundary stocks’ distribution across the 280 EEZs of 198 coastal countries/political entities under a high greenhouse gas emissions scenario (Methods). We then estimate the year in which the 633 transboundary species (9,132 stocks) are projected to shift their shared distribution beyond historical variations (Fig. \@ref(fig:toeFigure1N)) and the intensity of this shift (i.e., time of emergence and stock share ratio; Methods). We treat each species in an EEZ as a single stock due to the lack of more spatially resolved data to delineate the boundary of a managed population [@PalaciosAbrantes:2020tv] and only consider shared stocks between neighboring EEZs (i.e., excluding the high seas).

@@

We coupled a mechanistic population dynamic model (i.e., the DBEM) with an Earth System Model (i.e., the Geophysical Fluid Dynamics Laboratory Earth system model, GFDL) to project the future global distribution of transboundary stocks under climate change (see the Supplementary Materials). In brief, the DBEM algorithm uses historical species distribution and catch, eco-physiology, habitat suitability and spatial population dynamics to project abundance and potential fisheries catches in a $0.5^o$ x $0.5^o$ grid-cell of the world [@Cheung:2010dt; @Cheung:2009ud; @Cheung:2016jd]. The DBEM was driven with ten different simulations of the GFDL ESM2M (i.e., a ten-ensemble member simulation) each of them representing a unique atmosphere and ocean state at the start of the simulation [@Dunne:2012bo; @Dunne:2013db; @Frolicher:2020cc]. The simulations were run under historical forcing until 2005, afterwards and until 2100, simulations followed a high greenhouse gas emissions scenario, Representative Concentration Pathway 8.5 (RCP 8.5) [@Riahi:2011dk].

We then developed a Transboundary Index to evaluate range shifts in the shared distribution of transboundary stocks. The index is based on changes in the distance of the stock’s abundance centroid and that of the EEZs that share a given stock (Fig. \@ref(fig:toeFigure1)A). We used the ten-ensemble-member approach to determine the historical internal variability of the stock (i.e., 1951 to 2005) and apply the concept of *time of emergence*  to determine a shift in a stock’s shared distribution (see the Supplementary Materials - *Calculating shifts in the shared distribution of transboundary stocks*). The time of emergence (i.e., the year a transboundary stocks shifted from its historical distribution) was defined as the first year the index was beyond two standard deviations from the historical internal variability for a 95% probability of "emergence" (See Supplemental Material - Equ. \@ref(eq:four)). We also relaxed the results by including a threshold of one standard deviation (i.e., representing a probability of 68% that the index has emerged).

Finally, we apply the Game Theory concept of *threat point* [See @Sumaila:2020tp; @Munro:1979dg for its application to fisheries] to quantify the intensity of changes in the shared distribution of transboundary stocks between neighboring EEZs (See Supplemental Material - *Quantifying the intensity of transboundary stocks’ range shift*). A threat point is defined as the minimum payoff a player would be willing to receive to cooperate in a game-theoretic model. Here, we define the minimum payoff as two standard deviations from the historical catch proportion (average of 1951 to 2005) that each EEZ had for each stock. This assumes that countries are adapted to the historical sharing, but no country would be willing to collaborate if the future catch proportion of the transboundary stock is lower than it has ever been [@PalaciosAbrantes:2020vt; @Sumaila:2020tp]. Finally, we computed changes in catch proportion that were below the threat point of neighboring EEZs sharing a stock for two time periods relative to the historical proportion: 2030 (2021 and 2040) matching the UN 2030 Agenda and a mid-century timeline (2040 to 2050).

@@

### Climate risk hotspots for transboundary fisheries management
